final figures for inclusion

### 2021-04-20 shared analysis files /data_wrangling_plot.Rmd
trying to reboot my thesis formatting and also updating final results models and figures

```{r, echo=FALSE, results = 'hide', message=FALSE}
## loading libraries 
pacman::p_load(metafor, tidyverse, viridis, ggsci, plotly,  mgcv)
#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
#library(orchaRd)

@@

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -20) 
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range %in% c(10))
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

@@

```{r, echo= FALSE, message=FALSE}
#simple random effects model 
simple_model <- rma.mv(yi, vi, data=dat_MA_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
simple_model
```

@@

```{r, echo = FALSE}
#forest plot
forest <- normalized %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
```
**Figure 1.**
Forest plot broken out by study and experiment and their observed effected sizes and sampling variances.

@@

```{r, echo= FALSE, message= FALSE}
#GAM APPROACH
#fit GAM
gam1<-gam(yi~te(mean_temp_constant,flux_range,k=7),data=normalized) # note, lots of decisions that I blew by here
summary(gam1)
#estimate values via gam
vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
pd.gam<-predict(gam1,newdata=vals)
surf<-data.frame(vals,value=pd.gam)
vertices <- data.frame(
  x = c(0, 0, 20),
  y = c(10, 0, 0), 
  id = c("a", "a", "a")
)
#plot values as a surface
```

```{r, echo= FALSE, message= FALSE}
#plot gam surface
ggplot()+
  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range, fill=value))+
  geom_contour(data = surf, 
               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
               colour="black", breaks=c(0),size=2)+
  #geom_contour_filled()
  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
  scale_fill_gradientn('Effect Size',
                       colors= viridis(100),
                       breaks=seq(-2,5,2),limits=c(-2,5),
                       guide = guide_colourbar(ticks.colour = gray(.2),
                                               ticks.linewidth = 1.,
                                               frame.colour = "black"))+
  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
  theme_bw()+ 
  theme(legend.position="bottom")+
  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")
```
**Figure 3.** Contour plot of GAM generated effects sizes overlayed with raw effect sizes from the extracted dataset. Black lines denote the boundary from negative to positive effect sizes. Grey triange denotes excluded contour that lacks any raw data.

@@

```{r}
#with interaction term 
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
full_rf_model

#separate model with thermal history trait as a mod 
full_rf_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
                          ~flux_range * mean_temp_constant + stressful, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
full_rf_thermal
#absolute yi model 
absolute_model<-rma.mv(abs(yi), vi, data=dat_MA_ES, mods = 
                          ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
absolute_model
```

@@

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
  geom_point()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by organization level 
          and fit with linear model")
```
**Figure 4.**
We can also see that there a difference in effect size in organization level responses across fluctuation range. Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
  geom_jitter()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by 
          whether or mean temperature was thermally stressful")
```
**Figure 5.** 
When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_jitter()+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("SMD across fluctuation ranges colored by response type")

```
**Figure 6.** We cab see that the ranges at which traits are measured is much wider than rates, however, there appears to a minorly negative trend in effect sizes as traits increase in fluctuation range. The opposite is true for rates. Great way to show that there is a lack of information on rates in the collected dataset.

@@

- How do all the different models sound? Does it make sense to have these different models?
- Do the figures match up with the story/questions associated with them?
- Jittering versus standard plotting for the regression plots?
_________________________________
Supplementary Plots/Code

>>>>

```{r}
dat_MA_ES %>%
  count(study_id)
```

```{r, echo=FALSE, results = 'hide', message=FALSE}
## loading libraries 
pacman::p_load(metafor, tidyverse, viridis, visreg)
#pacman::p_load(ggsci, plotly,  mgcv)
#devtools::install_github("itchyshin/orchard_plot", subdir = "orchaRd", force = TRUE,build_vignettes = TRUE)
#library(orchaRd)

@@

#removing outlier point to get better picture of data structure AND excluding CT assays 
normalized <- dat_MA_ES %>%
  filter(yi > -4) 
#filtering for just flux ranges of 10 C 
common_range <- dat_MA_ES %>% 
  filter(flux_range %in% c(10))
#filtering for just studies with mean temp of 24
common_temp <- dat_MA_ES %>% 
  filter(mean_temp_constant == "24")
```

@@

```{r, echo= FALSE, message=FALSE}
#simple random effects model 
simple_model <- rma.mv(yi, vi, data=dat_MA_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
simple_model

absolute_model <- rma.mv(abs(yi), vi, data=dat_MA_ES, 
               random = ~1 | study_id / experiment_id/ response_id,
                 method="REML") 
absolute_model
```

@@

```{r, echo = FALSE}
#forest plot
forest <- dat_MA_ES %>%
  group_by(study_id, experiment_id) %>%
  summarise(mean_yi = mean(yi),
            mean_vi = mean(vi)) %>%
  unite(group, c(study_id, experiment_id), sep = " experiment ", remove = FALSE)
forest(forest$mean_yi, forest$mean_vi, slab=forest$group, pch =19, psize=1, col = "gray2")
```
**Figure 1.**
Forest plot broken out by study and experiment and their observed effected sizes and sampling variances.

@@

```{r, echo= FALSE, message= FALSE}
normalized %>%
ggplot(aes(fill=mean_temp_constant, y = yi, x=flux_range))+
  geom_jitter( alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Mean Temperature (C)', option = "C")+
  theme_classic()+
  xlab("Temperature range")+
  ylab("Effect Size")+
  geom_line(y=0, linetype = "dashed")+
  theme(legend.position="bottom")
```

```{r}
# keep this to justify range of (-4:5)
dat_MA_ES %>%
filter(yi > -20) %>%
  ggplot(aes(x=yi))+
  geom_histogram(binwidth = 1)
```
**Figure 3.** Contour plot of GAM generated effects sizes overlayed with raw effect sizes from the extracted dataset. Black lines denote the boundary from negative to positive effect sizes. Grey triange denotes excluded contour that lacks any raw data.

@@

```{r}
#with interaction term 
full_rf_model<-rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
full_rf_model
```

```{r}
#plot to include to demonstrate model estimates
full_model_coefficients<- as.data.frame(coef(full_rf_model)) %>%
  tibble::rownames_to_column("covariates")

full_model_coefficients %>%
  ggplot(aes(x = fct_reorder(covariates, coef(full_rf_model)),
             y = coef(full_rf_model),
             fill = coef(full_rf_model) > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(x = NULL, y = "Coefficient estimate") +
  scale_fill_viridis(discrete = TRUE)+
  theme_bw()
```

```{r}
#thermal history only bar graph

#separate model with thermal history trait as a mod 
only_thermal<-rma.mv(yi, vi, data=dat_MA_ES, mods = 
                          ~ stressful, 
               random = ~1 | experiment_id/ study_id/ response_id,
                 method="REML") 
only_thermal

thermal_only_coefficients<- as.data.frame(coef(only_thermal)) %>%
  tibble::rownames_to_column("covariates")

@@

```{r}
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(exp_age)))+
  geom_point()+
  labs(title = "SMD across fluctuation ranges colored by organism size and fit with linear model", 
       x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Experimental age")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("larval", "juvenile", "adult"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r}
# scatterplot of standardized mean response vs flux range colored and lm fit by organism size
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(size)))+
  geom_point()+
  labs(title = "SMD across fluctuation ranges colored by organism size and fit with linear model", 
       x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Size")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("extra-small", "small", "medium", "large"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by org level
ggplot(normalized, aes(x=flux_range, y=yi, color = factor(org_level)))+
  geom_point()+
  labs(title = "SMD across fluctuation ranges colored by organization level and fit with linear model", 
       x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Biological organization level")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("organismal", "population"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom") 
```
**Figure 4.**
We can also see that there a difference in effect size in organization level responses across fluctuation range. Population level reponses trended positively across fluctuation ranges, while organism level responses trended negatively across fluctuation ranges.

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
  drop_na(stressful) %>%
ggplot(aes(x=flux_range, y=yi, color = factor(stressful)))+
  geom_jitter()+
  labs(title = "SMD across fluctuation ranges colored by thermal stress experienced", 
       x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Thermally stressfull")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("no", "yes"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```
**Figure 5.** 
When we consider where the temperatures in the fluctuation range reach thermally stressful levels, there is a small difference between responses to stressful  temperatures (negative) and non-stressful temperatures (positive).

```{r, echo= FALSE}
# scatterplot of standardized mean response vs flux range colored and lm fit by life history thermal characteristic
normalized %>%
ggplot(aes(x=flux_range, y=yi, color = resp_type))+
  geom_jitter()+
  labs(title = "SMD across fluctuation ranges colored by response type", 
       x = "Temperature fluctuation range (C)",
       y ="SMD",
       color = "Life history metric")+
  scale_color_viridis(discrete = TRUE, option = "D", labels = c("rate", "trait"))+
  geom_smooth(method="lm", formula = y~x)+
  theme_bw()+
  theme(legend.position="bottom")
```
**Figure 6.** We cab see that the ranges at which traits are measured is much wider than rates, however, there appears to a minorly negative trend in effect sizes as traits increase in fluctuation range. The opposite is true for rates. Great way to show that there is a lack of information on rates in the collected dataset.

@@

- How do all the different models sound? Does it make sense to have these different models?
- Do the figures match up with the story/questions associated with them?
- Jittering versus standard plotting for the regression plots?

_________________________________
Supplementary Plots/Code

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
za <- dat_MA_ES %>%
  filter(study_id != "garcaruiz2011")

full_rf_thermal_2 <- rma.mv(yi, vi, data=dat_MA_ES, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type + stressful, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
full_rf_thermal_2
za_model <- rma.mv(yi, vi, data=za, mods = ~flux_range * mean_temp_constant +
                        exp_age + size + org_level + resp_type, 
               random = ~1 |  study_id/ experiment_id/ response_id,
                 method="REML") 
za_model
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#GAM APPROACH
#fit GAM
gam1<-gam(yi~te(mean_temp_constant + te(as.factor(flux_range), k=7)),data=normalized) # note, lots of decisions that I blew by here
summary(gam1)
#estimate values via gam
vals<-expand.grid(flux_range=seq(0,20,length.out = 100), mean_temp_constant=seq(0,35,length.out = 100))
pd.gam<-predict(gam1,newdata=vals)
surf<-data.frame(vals,value=pd.gam)
vertices <- data.frame(
  x = c(0, 0, 20),
  y = c(10, 0, 0), 
  id = c("a", "a", "a")
)
#plot values as a surface
```

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
#plot gam surface
ggplot()+
  geom_raster(data = surf, aes(x=mean_temp_constant,y=flux_range))+
  geom_contour(data = surf, 
               aes(x=mean_temp_constant,y=flux_range,z=value), colour=gray(0.2),binwidth=0.5,size=0.2)+
  geom_contour(data = surf,  aes(x=mean_temp_constant,y=flux_range,z=value), 
               colour="black", breaks=c(0),size=2)+
  #geom_contour_filled()
  scale_x_continuous('Mean Temp (°C)',expand=c(0,0))+
  scale_y_continuous('Temp Range (°C)',expand=c(0,0))+
  scale_fill_gradientn('Effect Size',
                       colors= viridis(100),
                       breaks=seq(-2,5,2),limits=c(-2,5),
                       guide = guide_colourbar(ticks.colour = gray(.2),
                                               ticks.linewidth = 1.,
                                               frame.colour = "black"))+
  coord_cartesian(xlim=c(0,35),ylim=c(0,20))+
  geom_jitter(data = normalized, mapping = aes(x=mean_temp_constant,y=flux_range,z= yi,fill = yi),
              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  scale_fill_viridis('Effect Size', limits = c(-2,5), option = "C")+
  theme_classic()+ 
  theme(legend.position="bottom")+
  geom_polygon(data = vertices, mapping = aes(x = x, y = y), fill = "gray")

```{r, echo= FALSE, results = 'hide', message=FALSE, fig.show='hide'}
yi_direcionality <- normalized %>%
  mutate(yi_direction = case_when(
    yi > 0 ~ "positive", 
    TRUE~ "negative"
  ))

yi_direcionality %>%
ggplot(aes(x=mean_temp_constant, y= flux_range, fill= as.factor(yi_direction)))+
  geom_jitter(data = yi_direcionality, 
              mapping = aes(x=mean_temp_constant,y=flux_range,fill = yi_direction),
              alpha = 0.9, pch=21,size=3,colour="black", width = 0.7, height = 0.7)+
  geom_smooth(method = "lm", color = "black")
```
